Title here

### 2017-01-04 portfolio/text/ms.Rmd
Edit to clarify

Individual revenue variability decreases exponentially with higher mean fished
species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
effort, strategies with one unit higher mean species diversity had on average `r
h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern
looks qualitatively similar with or without individuals targeting salmon. As
examples, variability was highest for the herring roe gillnet and king crab
fisheries. Most strategies with the lowest estimated revenue variability
combined halibut with a salmon permit and had high levels of fished species
diversity (Fig. \ref{fig:across}).

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue but also increased variability (Fig.
\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a permit set tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [REFS]. However, in practice, the
ability of individuals to effectively diversify is limited by regulation and
market forces. Here we document the effect of individual-level changes in
species diversity on revenue and revenue variability across three scales. We
find that (1) individuals with fishing permit sets that allow high levels of
species diversity tend to have lower revenue variability. (2) Adding a fisheries
permit usually increases revenue and decreases revenue variability. However, (3)
within the constraints of a permit and accounting for effort, specializing tends
to increase revenue while both specializing and generalizing tend to correlate
with greater revenue variability with considerable differences amongst
fisheries. Thus, the most effective mechanism by which species diversity can
lower revenue variability for individual fishers --- switching or adding fishing
permits --- is costly, limited by regulations, and consequently unavailable to
many individuals.

Our results across fishing strategies are in line with previous findings at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
results with respect to individual year-to-year behaviour provide new insights.
Within strategies, specializing tended to be associated with increased revenue.
This could be for a number of reasons. First, specialization can increase
fishing efficiency [@smith1986]. Efficiency gains from specialization can be
seen in other natural resource sectors. For example, rice farmers in Ghana can
increase their efficiency by devoting more land and labor to rice production
compared to other crops [@abdulai2000]. Second, given the availability of
valuable resources, specializing can be highly profitable. For instance, when
cocoa prices have been high, Nigerian farmers have tended to specialize on the
crop [@mustapha1999] and specializing on valuable tobacco farming can be the
optimal strategy regardless of alternative crops [@chavez2014]. In our analyses,
fishers with the opportunity to fish highly valuable sockeye salmon benefited
the most from specializing. However, specializing can have longterm risks if
market, biological, or environmental conditions change. For example, U.S.
farmers specialized on wheat, corn, or soybeans were hit harder than diversified
farmers in the early 1980s when a financial crisis hit [@ikerd1990]. Third,
generalization --- a lack of specialization --- can happen as a consequence of
non-ideal circumstances such as poor availability of a more valuable fish stock
in a given year. For example, downward economic shocks have been associated with
increased diversity for households relying on forests for income [@debela2012].

We also found that altering species diversity tended to be associated with
increased revenue variability. There are number of possible reasons for this
relationship. For instance, changes to fishing patterns will sometimes be
associated with a learning or adaptation phase as individuals become accustomed
to new locations, new gear, or new species [REF]. Additionally, changes in
diversity might reflect years in which targeted species were less available and
fishers were forced to target other species or landed more bycatch, adding
stochasticity to their revenue. Most importantly, this increased variability
reflects an inability for our large-scale model to predict an individual's
revenue in a subsequent year. There are many reasons, often unique to an
individual permit holder's behaviour [@hilborn1985] and therefore not included
in our large-scale model, that might make that specific individual's revenue
less predictable if species diversity changes.

@@

<!-- TODO: get fig:re in results -->
<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects -->
<!-- that vary by permit set (Fig. S\ref{fig:re}). -->

@@

\noindent This approach, combined with the lack of a global intercept,
constrains the mean change in log revenue across all fishers and years to be
zero. I.e., we assume that the data are stationary after accounting for all
covariates and including an intercept that varies with strategy-year
combinations. Initial versions of the model confirmed that the global intercept
along with strategy-level intercepts would be estimated at almost exactly zero.
The estimated $\beta_{j,t}$'s represent years when people with a particular
permit set tended to collectively do better or worse than the long-term average
(Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
(and variability) is after accounting for the general trend in revenue for that
permit set.

>>>>

Individual revenue variability decreases exponentially with higher mean fished
species diversity across combinations of state and federal fishing permits
("strategies"; Fig. \ref{fig:across}). Controlling for effort, strategies with
one unit higher mean species diversity had on average `r h1[[2]]`% (`r
h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern looks
qualitatively similar with or without individuals targeting salmon. As examples,
variability was highest for the herring roe gillnet and king crab fisheries.
Most strategies with the lowest estimated revenue variability combined halibut
with a salmon permit and had high levels of fished species diversity (Fig.
\ref{fig:across}).

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given set of permits is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort, specializing tends
to be associated with increased revenue but also increased variability (Fig.
\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a set of permits tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [REFS]. However, in practice, the
ability of individuals to effectively diversify is limited by regulation and
market forces. Here we document the effect of individual-level changes in
species diversity on revenue and revenue variability across three scales. We
find that (1) individuals with sets of permits that allow high levels of species
diversity tend to have lower revenue variability. (2) Adding a fisheries permit
usually increases revenue and decreases revenue variability. However, (3) within
the constraints of a permit and accounting for effort, specializing tends to
increase revenue while both specializing and generalizing tend to correlate with
greater revenue variability with considerable differences amongst fisheries.
Thus, the most effective mechanism by which species diversity can lower revenue
variability for individual fishers --- switching or adding fishing permits ---
is costly, limited by regulations, and consequently unavailable to many
individuals.

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
findings within strategies, with respect to individual year-to-year behaviour,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, biological, or environmental
conditions change. For example, U.S. farmers specialized on wheat, corn, or
soybeans were hit harder than diversified farmers in the early 1980s when
a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource in a given year. For example,
downward economic shocks have been associated with increased diversity for
households relying on forests for income [@debela2012].

We also found that, within strategies, altering species diversity tended to be
associated with increased revenue variability. There are number of possible
reasons for this relationship. For instance, changes to fishing patterns will
sometimes be associated with a learning or adaptation phase as individuals
become accustomed to new locations, new gear, or new species [REF].
Additionally, changes in diversity might reflect years in which targeted species
were less available and fishers were forced to target other species or landed
more bycatch, adding stochasticity to their revenue. Most importantly, this
increased variability reflects an inability for our large-scale model to predict
an individual's revenue in a subsequent year. There are many reasons, often
unique to an individual permit holder's behaviour [@hilborn1985] and therefore
not included in our large-scale model, that might make a specific individual's
revenue less predictable if species diversity changes.

@@

<!-- TODO: get fig:re in results -->
<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and
effects --> <!-- that vary by strategy (Fig. S\ref{fig:re}). -->

@@

\noindent This approach, combined with the lack of a global intercept,
constrains the mean change in log revenue across all fishers and years to be
zero. I.e., we assume that the data are stationary after accounting for all
covariates and including an intercept that varies with strategy-year
combinations. Initial versions of the model confirmed that the global intercept
along with strategy-level intercepts would be estimated at almost exactly zero.
The estimated $\beta_{j,t}$'s represent years when people with a particular set
of permits tended to collectively do better or worse than the long-term average
(Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
(and variability) is after accounting for the general trend in revenue for that
set of permits.
